### 2023-01-24 latex/again.tex
converted to peerj overleaf template

\hypertarget{sec:introduction}{%
\section{Introduction}\label{sec:introduction}}

@@

\hypertarget{sec:background}{%
\section{Background and related work}\label{sec:background}}

@@

\hypertarget{sec:fdo}{%
\subsection{FAIR Digital Object}\label{sec:fdo}}

\hypertarget{sec:next-step-fdo}{%
\subsubsection{Next steps for FDO}\label{sec:next-step-fdo}}

@@

\hypertarget{fdo-approaches}{%
\subsubsection{FDO approaches}\label{fdo-approaches}}

@@

\hypertarget{sec:ld}{%
\subsection{From the Semantic Web to Linked Data}\label{sec:ld}}

@@

\hypertarget{sec:semweb}{%
\subsubsection{A brief history of the Semantic Web}\label{sec:semweb}}

@@

\hypertarget{sec:ld-web}{%
\subsubsection{Linked Data: Rebuilding the Web of Data}\label{sec:ld-web}}

@@

\hypertarget{sec:comparing}{%
\section{Comparing FDO and existing approaches}\label{sec:comparing}}

@@

\hypertarget{sec:interoperability-compare}{%
\subsection{Considering FDO/Web as interoperability framework for Fast Data}\label{sec:interoperability-compare}}

@@

\hypertarget{mapping-of-metamodel-concepts}{%
\subsubsection{Mapping of Metamodel concepts}\label{mapping-of-metamodel-concepts}}

@@

\hypertarget{sec:doip-fdo-compare}{%
\subsection{Assessing FDO implementations}\label{sec:doip-fdo-compare}}

@@

\begin{itemize}
\tightlist
\item
  G1 and G2 call for stability and trustworthiness. While the foundations of both DOIP and Linked Data approaches are now well established -- the FDO requirements and in particular how they can be implemented are still taking shape and subject to change.
\item
  Machine actionability (G4, G6) is a core feature of both FDOs and Linked Data. Conceptually they differ in the which way types and operations are discovered, with FDO seemingly more rigorous. In practice, however, we see that DOIP also relies on dynamic discovery of operations and that operation expectations for types (FDOF7) have not yet been defined.
\item
  FDO proposes that types can have additional operations beyond CRUD (FDOF5, FDOF6), while Linked Data mainly achieves this with RESTful patterns using CRUD on additional resources, e.g.~\texttt{order/152/items}. These are mainly stylistics but affect the architectural view -- FDOs have more of an object-oriented approach.
\item
  FDO puts strong emphasis on the use of PIDs (FDOF1, FDOF2, FDOF3, FDOF5), but in current practice DOIP use local types, local extended operations (FDOF5) and attributes (FDOF4) that are not bound to any global namespace.
\item
  Linked Data have a strong emphasis on semantics (FDOF8), and metadata schemas developed by community agreements (FDOF10). FDO types need to be made reusable across servers.
\item
  While FDO recommends nested metadata FDOs (FDOF8, FDOF9), in practice this is not found (or linked with custom keys), particularly due to lack of namespaces and the favouring of local types rather than type/property re-use. Linked Data frequently have multiple representations, but often not sufficiently linked, perhaps \texttt{prov:specializationOf} (\protect\hyperlink{ref-9T4j3N4e}{{`PROV-O: The PROV Ontology'}})
\item
  FDO collections are not yet defined for DOIP, while Linked Data seemingly have too many alternatives, LDP has specific native support for containers.
\item
  Tombstones for deleted resources are not well supported, nor specified, for either approach, although the continued availability of metadata when data is removed is a requirement for FAIR principles (see RDA-A2-01M in table \ref{sec:fair-compare}).
\item
  DOIP supports multiple chunks of data for an object (FDOF3), while Linked Data can support content-negotiation. In either case it can be unclear to clients what is the meaning or equivalence of any additional chunks.
\end{itemize}

\hypertarget{sec:middleware}{%
\subsection{Comparing FDO and Web as middleware infrastructures}\label{sec:middleware}}

@@

\hypertarget{sec:fair-compare}{%
\subsection{Assessing FDO against FAIR}\label{sec:fair-compare}}

@@

\hypertarget{eosc-interoperability-framework}{%
\section{EOSC Interoperability Framework}\label{eosc-interoperability-framework}}

@@

\hypertarget{sec:discussion}{%
\section{Discussion}\label{sec:discussion}}

@@

\hypertarget{what-does-it-mean-for-linked-data}{%
\subsection{(What does it mean for Linked Data?)}\label{what-does-it-mean-for-linked-data}}

@@

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

@@

\hypertarget{references}{%
\section{References}\label{references}}

>>>>

\hypertarget{sec:introduction}{%
\section*{Introduction}\label{sec:introduction}}

@@

\hypertarget{sec:background}{%
\section*{Background and related work}\label{sec:background}}

@@

\hypertarget{sec:fdo}{%
\subsection*{FAIR Digital Object}\label{sec:fdo}}

\hypertarget{sec:next-step-fdo}{%
\subsubsection*{Next steps for FDO}\label{sec:next-step-fdo}}

@@

\hypertarget{fdo-approaches}{%
\subsubsection*{FDO approaches}\label{fdo-approaches}}

@@

\hypertarget{sec:ld}{%
\subsection*{From the Semantic Web to Linked Data}\label{sec:ld}}

@@

\hypertarget{sec:semweb}{%
\subsubsection*{A brief history of the Semantic Web}\label{sec:semweb}}

@@

\hypertarget{sec:ld-web}{%
\subsubsection*{Linked Data: Rebuilding the Web of Data}\label{sec:ld-web}}

@@

\hypertarget{sec:comparing}{%
\section*{Comparing FDO and existing approaches}\label{sec:comparing}}

@@

\hypertarget{sec:interoperability-compare}{%
\subsection*{Considering FDO/Web as interoperability framework for Fast Data}\label{sec:interoperability-compare}}

@@

\hypertarget{mapping-of-metamodel-concepts}{%
\subsubsection*{Mapping of Metamodel concepts}\label{mapping-of-metamodel-concepts}}

@@

\hypertarget{sec:doip-fdo-compare}{%
\subsection*{Assessing FDO implementations}\label{sec:doip-fdo-compare}}

@@

\begin{itemize}
\tightlist
\item
  G1 and G2 call for stability and trustworthiness. WhileC the foundations of both DOIP and Linked Data approaches are now well established -- the FDO requirements and in particular how they can be implemented are still taking shape and subject to change.
\item
  Machine actionability (G4, G6) is a core feature of both FDOs and Linked Data. Conceptually they differ in the which way types and operations are discovered, with FDO seemingly more rigorous. In practice, however, we see that DOIP also relies on dynamic discovery of operations and that operation expectations for types (FDOF7) have not yet been defined.
\item
  FDO proposes that types can have additional operations beyond CRUD (FDOF5, FDOF6), while Linked Data mainly achieves this with RESTful patterns using CRUD on additional resources, e.g.~\texttt{order/152/items}. These are mainly stylistics but affect the architectural view -- FDOs have more of an object-oriented approach.
\item
  FDO puts strong emphasis on the use of PIDs (FDOF1, FDOF2, FDOF3, FDOF5), but in current practice DOIP use local types, local extended operations (FDOF5) and attributes (FDOF4) that are not bound to any global namespace.
\item
  Linked Data have a strong emphasis on semantics (FDOF8), and metadata schemas developed by community agreements (FDOF10). FDO types need to be made reusable across servers.
\item
  While FDO recommends nested metadata FDOs (FDOF8, FDOF9), in practice this is not found (or linked with custom keys), particularly due to lack of namespaces and the favouring of local types rather than type/property re-uCse. Linked Data frequently have multiple representations, but often not sufficiently linked, perhaps \texttt{prov:specializationOf} (\protect\hyperlink{ref-9T4j3N4e}{{`PROV-O: The PROV Ontology'}})
\item
  FDO collections are not yet defined for DOIP, while Linked Data seemingly have too many alternatives, LDP has specific native support for containers.
\item
  Tombstones for deleted resources are not well supported, nor specified, for either approach, although the continued availability of metadata when data is removed is a requirement for FAIR principles (see RDA-A2-01M in table \ref{sec:fair-compare}).
\item
  DOIP supports multiple chunks of data for an object (FDOF3), while Linked Data can support content-negotiation. In either case it can be unclear to clients what is the meaning or equivalence of any additional chunks.
\end{itemize}

\hypertarget{sec:middleware}{%
\subsection*{Comparing FDO and Web as middleware infrastructures}\label{sec:middleware}}

@@

\hypertarget{sec:fair-compare}{%
\subsection*{Assessing FDO against FAIR}\label{sec:fair-compare}}

@@

\hypertarget{eosc-interoperability-framework}{%
\section*{EOSC Interoperability Framework}\label{eosc-interoperability-framework}}

@@

\hypertarget{sec:discussion}{%
\section*{Discussion}\label{sec:discussion}}

@@

\hypertarget{what-does-it-mean-for-linked-data}{%
\subsection*{(What does it mean for Linked Data?)}\label{what-does-it-mean-for-linked-data}}

@@

\hypertarget{conclusion}{%
\section*{Conclusion}\label{conclusion}}

@@

\hypertarget{references}{%
\section*{References}\label{references}}
